---
output: 
  officedown::rdocx_document:
    reference_docx: styles.docx
params:
  table: "granddata.d36109"
  year: 2020
  pollutant: 98
  report_type: "demo"
  # graph: "multiline"
  graph: "donut"
  # set: "fueltype"
  set: "sourcetype"
always_allow_html: true
---

```{r setup, include = FALSE}
# Set options
knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE)
# Load required packages
library(officer) # need to load this from source
library(officedown) # making docx out of Rmd files.
library(httr) # API queries
library(jsonlite) # working with jsons
library(dplyr) # data wrangling
#library(stringr, lib.loc = "../lib") # String manipulation
#library(magrittr, lib.loc = "../lib") 
library(ggtext) # adding HTML into data visualizations
# library(ggplot2) # data visualization
library(catviz) # for making visualizations
library(stringr) # for string manipulation
library(purrr) # for tidy for loops

retrieve_p = function(table = "granddata.36109", year = 2025, pollutant = 98){
  # table = "granddata.d36109"; year = 2025; pollutant = 98;
  
  # Example Query:
  # https://api.cat-apps.com/visualizer/v1/retrieve_p/?table=granddata.d36109&year=2025&pollutant=98

  base = "https://api.cat-apps.com/visualizer/v1/retrieve_p/"
  # Make a table id
  url = paste0(base, "?", "table=", table, "&",
               "year=", year, "&", 
               "pollutant=", pollutant)
  
    # Get headers  
  headers = httr::add_headers("Content-Type" = "application/json")

  # Send query
  result = httr::GET(url = url, headers)

  # If the query cannot be completed, return NULL
  if(result$status_code != 200){ return(NULL)  }

  # Otherise, return this and process it into an R list object
  output = result %>%
    with(content) %>%
    rawToChar() %>%
    jsonlite::fromJSON()

  return(output)
}

# Write a data-retrieval function
retrieve_data = function(geoid = "36109", year = 2025, pollutant = 98, set = "sourcetype", metric = "emissions"){
  
  # Testing values
  # geoid = "36109"; year = 2025; pollutant = 98; set = "sourcetype"; metric = "emissions"
  
  base = "https://api.cat-apps.com/reporter/v1/retrieve_data/donut"
  url = paste0(base, "?", "geoid=", geoid, "&",
               "year=", year, "&", 
               "pollutant=", pollutant, "&",
               "set=", set, "&",
               "metric=", metric)

  # Goal: this kind of API query
  # https://api.cat-apps.com/reporter/v1/retrieve_data/donut?geoid=36109&year=2025&pollutant=98&set=fueltype&metric=vmt

  # Get headers  
  headers = httr::add_headers("Content-Type" = "application/json")

  # Send query
  result = httr::GET(url = url, headers)

  # If the query cannot be completed, return NULL
  if(result$status_code != 200){ return(NULL)  }

  # Otherise, return this and process it into an R list object
  output = result %>%
    with(content) %>%
    rawToChar() %>%
    jsonlite::fromJSON()

  return(output)

}

# Apply the functions and retrieve data from CAT API.
p = retrieve_p(table = params$table, year = params$year, pollutant = params$pollutant)

# Access the geoid of that table
# geoid = params$table %>% 
#   str_extract("(d[0-9]{5}|d[0-9]{2})") %>% 
#   str_remove("d") 

# Generate the tabulated data for a bar / donut plot.
tab = retrieve_data(geoid = p$poi$geoid, year = params$year, pollutant = params$pollutant, set = params$set, metric = "emissions")

# Generate the visual
if(params$graph == "donut"){
  viz = catviz::visualize_donut(tab = tab, plotly = FALSE)  
}else if(params$graph == "bar"){
  viz = catviz::visualize_bar(tab = tab, plotly = FALSE)  
}else{ stop("That plot type is not currently supported.") }


# Hidden code chunk for background calculations.
# Make a pluralizer function
pluralize = function(word){
  # Make type plural if need be
  # Get last character
  letter = stringr::str_sub(word, -1, -1) 
  # Make it plural based on some rules
  pluralized = if(letter == "s"){ 
    "ses"
  }else if(letter == "y"){
    "ies"    
  }else{ paste0(letter, "s") }
  # Overwrite label with type
  word = word %>% str_sub(1, -2) %>% paste0(pluralized)  
  return(word)
}

bullet = function(tab, i){
  # Testing values
  # tab; i = 1
  
  # Get data from tab object  
  output = tab$data %>% 
    # Get ith row
    dplyr::slice(i) %>%
    # Get these variables
    dplyr::select(type, year, label_value, label_percent) 
  # Pluralize type
  output$type = pluralize(output$type)
  # Return it
  return(output)
}

capitalize_first= function(words){
  word_list=strsplit(words, split = " ")
  result=''
  for (i in 1:length(word_list[[1]])){
    word=word_list[[1]][i]
    if (i>1){
      word=tolower(word)
    }
    result=paste(result,word,sep=" ")
  }
  return(result)
}


# Check how many rows / slices are in my donut?
# n = tab %>% nrow()
b1 = bullet(tab = tab, i = 1)
# b1$type
# b1$year
# b1$label_value
b2 = bullet(tab = tab, i = 2)

# Make bullet information
n = tab$.n
b3 = 3:n %>% purrr::map_dfr(~bullet(tab = tab, i = .))
b3$title = paste0(b3$type[1], ", ", b3$type[2], ", and ", b3$type[3])

# Find a way to reference that set type
set_text = switch(
  EXPR = params$set,
  "sourcetype" = "types of vehicles",
  "fueltype" = "types of fuel",
  "regclass" = "regulatory classes",
  "roadtype" = "types of roads",
  "types")

# Find a way to reference that set type
set_singular = switch(
  EXPR = params$set,
  "sourcetype" = "type of vehicle",
  "fueltype" = "type of fuel",
  "regclass" = "regulatory class",
  "roadtype" = "type of road",
  "type")

# switch(
#   EXPR = params$set,
#   "sourcetype" = "vehicles",
#   "fueltype" = "fuels",
#   "regclass" = "regulatory classes",
#   "roadtype" = "road types",
#   "types")

```

```{r logo, echo=FALSE}
# LOGO
section_logo <- fpar(
  external_img(src = "logo_cat.png", height = 3/5, width = 6/5, alt = "CAT Logo"),
  ftext("                                                        ", fp_text(font.size = 15)),
  external_img(src = "logo_reporter.png", height = 3/5, width = 10/5, alt = "REPORTER Logo"),
  fp_p = fp_par(text.align = "left", keep_with_next = TRUE)
)

# Output logo
section_logo
```

```{r header, echo=FALSE}
# HEADER

# Example
# statement = "CO2e Emissions Distribution in Tompkins County 2020"
pollutant_simple = if_else(
  condition = p$pollutant_label == "CO2e",
  true = "Carbon", false = p$pollutant_label)

statement = paste0(pollutant_simple, " ", "Emissions in ", 
                   p$poi$name, " ", p$poi$area, ", ", p$year) 

# Make the header
bold_face <- shortcuts$fp_bold(font.size = 20)
bold_greenface <- update(bold_face, color = "#648fff")
section_header <- fpar(
  ftext(statement, prop = bold_greenface ),
  run_linebreak(),
  ftext("Made with CAT VISUALIZER by Gao Labs @ Cornell University."),
  fp_p = fp_par(text.align = "center"))

# Output header
section_header
```

New Research Shed Light on the Impact of Vehicle Emissions on `r p$pollutant_label` and `r b1$type`, `r b2$type`, `r b3$type[1]`, `r b3$type[2]`, `r b3$type[3]`. Estimates from the Climate Action in Transportation (CAT) Dashboard from Cornell University are empowering communities with vital data for a healthier and more sustainable environment.

<!---BLOCK_MULTICOL_START--->

```{r echo=FALSE, message=FALSE, fig.keep='all',dpi=300,fig.width=4, fig.height=4, fig.align = 'left'}
# Render plot
viz
# knitr::include_graphics(path = g$plot_path)
```

`r run_columnbreak()`

1. **`r b1$type` Take the Lead:** `r b1$type` are the primary culprits, accounting for a staggering `r b1$label_percent` of total emissions. This indicates that they are either more abundant in number, have higher emission rates, or both. \
2. **`r b2$type`, Still a Concern:** Surprisingly, combined emissions from `r tolower(b2$type)` come in second at `r b2$label_percent`. This demonstrates the collective impact of these `r set_text` in the region.\
3. **`r b3$title[1]`:** `r capitalize_first(b3$type[1])` contribute `r b3$label_percent[1]` of the `r p$pollutant_label` emissions, followed by `r tolower(b3$type[2])` at `r b3$label_percent[2]`. `r b3$type[3]` represent a mere `r b3$label_percent[3]`.



<!---BLOCK_MULTICOL_STOP{widths: [4,3], space: 0.1, sep: true}--->

In conclusion, while each `r set_singular` contributes to the overall `r tolower(pollutant_simple)` emissions of `r paste0(p$poi$name, " ", p$poi$area)`, efforts to reduce emissions might be most effective if they initially focus on `r tolower(b1$type)` and `r tolower(b2$type)`. It provides a clearer picture of where interventions could be targeted for a low-emissions future.

```{r echo=FALSE}
section_footer <- fpar(
  ftext("Contact Information:"),
  run_linebreak(),
  ftext("Department Name, Organization Name"),
  run_linebreak(),
  ftext("Street Address, City, State"),
  run_linebreak(),
  ftext("Email: name@thing.com"),
  fp_p = fp_par(text.align = "right"))

section_footer
```

